Exploring Bounds

### 2015-04-19 draft.tex
compared to greismer
getting there

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfig}

@@

\newtheorem{theorem}{Theorem}
%\numberwithin{theorem}{subsection}

@@

@@

@@

\begin{figure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{plotkin_q2.png}
  \caption{1a}
  \label{fig:sfig1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{plotkin_q3.png}
  \caption{1b}
  \label{fig:sfig2}
\end{subfigure}
\caption{plots of....}
\label{fig:fig}
\end{figure}

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfig}
\usepackage{enumitem}

@@

\newtheorem{theorem}{Theorem}
\numberwithin{theorem}{subsection}

@@

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\numberwithin{defn}{subsection}

\theoremstyle{definition}
\newtheorem{claim}{Claim}
\numberwithin{claim}{subsection}

@@

\begin{comment}
PLAN
do elias, don't prove first lemma, prove second lemma and thm
do given examples, draw graphs
finish out examples of lexicodes
add graphs to plotkin
finish analysis of asymptotic singleton and plotkin
\end{comment}

@@

\begin{comment}
\begin{figure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{plotkin_q2.png}
  \caption{1a}
  \label{fig:sfig1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.8\linewidth]{plotkin_q3.png}
  \caption{1b}
  \label{fig:sfig2}
\end{subfigure}
\caption{plots of....}
\label{fig:fig}
\end{figure}
\end{comment}

\section{Lexicodes}
We end our discussion on bounds with an interesting class of binary linear codes
called lexicodes, short for lexicographic codes.  Lexicodes are linear and meet the Gilbert bound; we will discuss the former and prove the latter.  Note that we are 
working in $\mathbb{F}_2$.

We construct the lexicode $\mathcal{L}$ of length $n$ and minimum distance $d$ using a basic greedy algorithm as follows.  First, order all $n$-tuples in lexicographic order:
$0\cdots000$, $0\cdots001$, $0\cdots010$, $0\cdots011$, $0\cdots100$, et cetera.  Put the zero vector $\mathbf{0}$ in $\mathcal{L}$.  Next, find the first vector $\mathbf{x}$
of weight $d$ in the lexicographic ordering, and put it in $\mathcal{L}$.  Then, find the next vector in  the lexicographic ordering whose distance from each vector in $\mathcal{L}$ is 
$d$ or more and add it to $\mathcal{L}$, repeating this process until you have used up your lexicographic list.

The set $\mathcal{L}$ is a linear code of length $n$ and minimum distance $d$.  The set $\mathcal{L}$ also contains linear subcodes which are generated by the above algorithm,
but stopping at the right spots.

\begin{theorem}
After constructing $\mathcal{L}$ as above, label the vectors $\mathbf{c}_0$, $\mathbf{c}_1,$ $\ldots$ in the order they were selected such that $\mathbf{c}_0$ is the $\mathbf{0}$ 
vector.

\begin{enumerate}
\item $\mathcal{L}$ is a linear code and the vectors $\mathbf{c}_{2^i}$ are a basis of $\mathcal{L}$.  
\item After $\mathbf{c}_{2^i}$ is chosen, the next $2^i -1$ vectors can be rewritten as $\mathbf{c}_1 + \mathbf{c}_{2^i}$, $\mathbf{c}_2 + \mathbf{c}_{2^i}$, $\ldots$, 
$\mathbf{c}_{2^i -1} + \mathbf{c}_{2^i}$.
\item Let $\mathcal{L}_i = \{\mathbf{c}_0,\mathbf{c}_1,\ldots,\mathbf{c}_{2^i -1}\}$.  Then $\mathcal{L}_i$ is an $[n,i,d]$ linear code.

\end{enumerate}

\end{theorem}

Though I won't prove this, it is not difficult to see that $1.$ and $3.$ follow from $2.$    In $2.$, we are showing that all codewords in the lexicode are simply linear combinations of 
the $c_{2^i}$ codewords, hence making the $c_{2^i}$ codewords a basis for the code.

The codes $\mathcal{L}_i$ satisfy the inclusions $\mathcal{L}_1 \subset \mathcal{L}_2 \subset \cdots \subset \mathcal{L}_k = \mathcal{L}$, where $k$ is the dimension of 
$\mathcal{L}$.  In general, this dimension value $k$ is not known until you have completed the construction as described above.

\begin{exmp}
Consider the codes $\mathcal{L}_i$ of length $5$ and minimum distance $2$.  First, we must construct the entire lexicode $\mathcal{L}$.  To do so, we enumerate
all 32 vectors of $\mathbb{F}_2^5$ in lexicographic order: $00000$,$00001$,$00010$,$\ldots$,$11111$.  We find that $\mathcal{L} = \{00000, 00011,00101,00110,01001,
01010,01100,01111,10001,10010,\\ 10100,10111,11000,11011,11101,11110\}$.  As a sanity check, note that $\left | \mathcal{L} \right | = 16$, a power of 2.  Consequently, we have
$\mathcal{L}_1 = \{00000,00011\}$, $\mathcal{L}_2 = \{00000, 00011,00101,\\ 00110\}$, $\mathcal{L}_3 = \{00000, 00011,00101,00110,01001,01010,01100,01111\}$, and
$\mathcal{L}_4 = \mathcal{L}$. We can check to make sure that each $\mathbf{c}_j$ is a linear combination of the $\mathbf{c}_{2^i}$s. Consider $\mathcal{L}_3$:

\begin{enumerate}[start=0]
\item $\mathbf{c}_0$ is the trivial linear combination.
\item $\mathbf{c}_1 = \mathbf{c}_1$
\item $\mathbf{c}_2 = \mathbf{c}_2$
\item $\mathbf{c}_3 = \mathbf{c}_1 + \mathbf{c}_2$.  $00110 = 00011+ 00101$.
\item $\mathbf{c}_4 = 1\cdot\mathbf{c}_4$
\item $\mathbf{c}_5 = \mathbf{c}_1 + \mathbf{c}_4$. $01010 = 00011 + 01001$.
\item $\mathbf{c}_6 = \mathbf{c}_2 + \mathbf{c}_4$. $011000 = 00101 + 01001$.
\item $\mathbf{c}_7 = \mathbf{c}_3 + \mathbf{c}_4 = \mathbf{c}_1 + \mathbf{c}_2 + \mathbf{c}_4$. $01111 = 00011 + 00101+01001$.
\end{enumerate}

The same analysis applies for all $\mathcal{L}_i$s.

Note that the greedy algorithm we descibed in the beginning of the section requires a strictly lexigraphical ordering, or else the code that is produced will not necessarily
be linear.  Again consider the case of $\mathbb{F}_2^5$, but instead of the lexigraphical ordering, shift every vector up one such that the ordering is identical except now the
zero vector is at the bottom of the list.  If we run our algorithm on this ordering, we will clearly not choose the zero vector to be a part of $\mathcal{L}$ and hence 
$\mathcal{L}$ will not be linear.

Also, note that lexicodes meet one of the bound we discussed in lecture, the Gilbert bound.  Recall the Gilbert bound:

\begin{equation}
B_q(n,d) \ge \frac{q^n}{\sum_{i=0}^{d-1} \binom{n}{i}(q-1)^{i}}
\end{equation}

\begin{claim}
Lexicodes meet the Gilbert bound.
\end{claim}

\begin{proof}
Consider an arbitrary lexicode $\mathcal{L}$.  Recall our construction of $\mathcal{L}$ with vectors in $\mathbb{F}_q^n$, wherein we visit each vector in our ordering,
and if it is distance $d$ or more to every current element in $\mathcal{L}$ then we add it to our set.  More importantly, if a vector is distance $d-1$ or less to any vector in
$\mathcal{L}$, then it is not selected.  Since we scan over every possible vector in $\mathbb{F}_q^n$, this implies that all vectors in $\mathbb{F}_q^n$ are distance less than
or equal to $d-1$ from any vector in $\mathcal{L}$, as if at any point of our algorithm we reached some $v \in \mathbb{F}_q^n$, $v \not\in \mathcal{L}$,  we would have then
added it $\mathcal{L}$, a contradiction.  Hence the covering radius of $\mathcal{L}$ is $d-1$.

Now, it is enough to show that any code $\mathcal{C}$ with covering radius $d-1$ or less meets the Gilbert bound.  Recall that the Gilbert bound can be thought of as a restatement of the 
following fact: (number of spheres around codewords) $\cdot$ (number of vectors in each sphere) $\ge$ (number of vectors in $\mathbb{F}_q^n$).  Since our covering radius
is $d-1$ or less, none of the spheres for our lexicodes will overlap, and thus there will be no double counting of vectors when counting up the vectors in each individual sphere.
Hence, $\mathcal{C}$ meets the Gilbert bound.
\end{proof}
